---
title: "Assignment 4 - Coordinating Heart Rate"
author: "Riccardo Fusaroli"
date: "November 6, 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(pastecs)
library(Metrics)
library(tseriesChaos)
library(SparseM)
library(crqa)
library(ggplot2)
library(tidyverse)
library(stringr)
library(readr)
library(lme4)
library(MuMIn)
library(lmerTest)
library(caret)
library(pROC)
library(dplyr)
library(groupdata2)
library(gridExtra)

wd =setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4")
```

## Analysing Heart Rate and Respiration data
The goal of this assignment is to first familiarize you with heart rate, and respiration data and their preprocessing. The second part explores how to analyze interpersonal coordination of these signals.

These are the questions you need to be able to answer at the end of the assignment (aka that you need to submit as part of the portfolio)

1) How do you preprocess heart rate and respiration data? Describe the process. If any data needs to be excluded, list the excluded data and motivate the exclusion.

2) Do you observe interpersonal coordination in heart rate and respiration? Describe your control baseline, the method used to quantify coordination, and the statistical models used to infer whether coordination was higher than in the baseline. Report the results of the models.

3) Do you observe differences in coordination between conditions? Report the models and results.

4) Is respiration coordination a likely driver of heart rate coordination? Describe how you would test for it. Bonus points if you actually run the tests and report methods and results.

N.B. to give you a bit more data I included data from last year (Study1) and from your class (Study2). Note that synchronouns and turn-taking are the same across both studies, but the third condition is different: last year it was self-paced joint reading; this year it was the tv-series conversation. So you might want to exclude the self-paced reading (but, up to you!)
```{r}
wd =setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4")

d1 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/Study2_G5_T1_Synchronous.csv")


plot1 =ggplot(data)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))

plot2 =ggplot(data)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))



#artifacts
removeOuts <- function(ts, threshold) {
  ts[ts > (mean(ts, na.rm = T) + (threshold * sd(ts, na.rm = T))) |
        ts < (mean(ts, na.rm = T)  - (threshold * sd(ts, na.rm = T)))] = mean(ts, na.rm =
  T)
return(ts)}

threshold=2.5
plot(d1$HR1)
d1$HR1=removeOuts(d1$HR1,threshold)
plot(d1$HR1)
plot(d1$HR2)
d1$HR2=removeOuts(d1$HR2,threshold)
plot(d1$HR2)
plot(d1$Resp1)
d1$Resp1 =removeOuts(d1$Resp1,threshold)
plot(d1$Resp1)
plot(d1$Resp2)
d1$Resp2 = removeOuts(d1$Resp2,threshold)
plot(d1$Resp2)

#downsampling
d1 = d1 %>% 
  group(n = 100, method = 'greedy') %>% 
  dplyr::summarise(time = mean(time,na.rm=T), HR1 = mean(HR1,na.rm=T),HR2 = mean(HR2,na.rm=T),Resp1 = mean(Resp1,na.rm=T),Resp2 = mean(Resp2,na.rm=T))

#scale
d1$Resp1S=scale(d1$Resp1)  
d1$Resp2S=scale(d1$Resp2)  
d1$HR1S=scale(d1$HR1)
d1$HR2S=scale(d1$HR2)

#plot
plot3 =ggplot(d1)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))

plot4 =ggplot(d1)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))

grid.arrange(plot1, plot2, plot3, plot4, nrow =2, ncol =2)

```

```{r}
#respiration
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")

ans = try(optimizeParam(d1$Resp1S, d1$Resp2S, par, min.rec = 3.5, max.rec = 4.5))

Results=crqa(d1$Resp1S,d1$Resp2S, delay=ans$delay, embed=ans$emddim, radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
Results

#heart
par = list(lgM =  50, steps = seq(1, 6, 1),  radiusspan = 100,  radiussample = 40, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")

ans = try(optimizeParam(d1$HR1S, d1$HR2S, par, min.rec = 3.5, max.rec = 4.5))

Results=crqa(d1$HR1S,d1$HR2S, delay=ans$delay, embed=ans$emddim, radius=ans$radius,normalize=0,rescale=0,mindiagline = 2,minvertline = 2)
Results
```



## Step by step suggestions to solve the assignment

### Exploring physiological signals

- Choose one pair (one pair, three conditions)
- Load the logs
- Produce a plot of the participants' respiration signal and a different one of the participants' HR signal (for inspecting whether the data is usable)
  N.B: remember the slides: artifacts, downsampling, scaling.
  N.B. The gridExtra::grid.arrange() function allows you to display the plots side by side. E.g. grid.arrange(plot1, plot2, plot3, ncol=3)
- Can you eye-ball which condition if any displays more physiological coordination?

- Run crqa on heart rate and respiration data (find parameters, run crqa)
- Does this tell you more than just eyeballing the plots?

```{r}

#scaling loop
files <- list.files(pattern = "csv")
files
# myfiles = lapply(files, read.delim)


for (i in files) {
  print(i)
  file = read.csv(i)
  #downsampling
  d1 = file %>%
  group(n = 100, method = 'greedy') %>%
  dplyr::summarise(
    time = mean(time, na.rm = T),
    HR1 = mean(HR1, na.rm = T),
    HR2 = mean(HR2, na.rm = T),
    Resp1 = mean(Resp1, na.rm = T),
    Resp2 = mean(Resp2, na.rm = T))
  #artifacts
  d1$HR1 = removeOuts(d1$HR1, threshold)
  d1$HR2 = removeOuts(d1$HR2, threshold)
  d1$Resp1 = removeOuts(d1$Resp1, threshold)
  d1$Resp2 = removeOuts(d1$Resp2, threshold)


  #scale
  d1$Resp1S = scale(d1$Resp1)
  d1$Resp2S = scale(d1$Resp2)
  d1$HR1S = scale(d1$HR1)
  d1$HR2S = scale(d1$HR2)
  
  #save in new folder
  file1 = paste0("pp_", i)
  write_csv(d1, file.path("pp", file1))
}



```
```{r}

setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp")
dp1 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G1_T1_Synchronous.csv")
dp2 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G1_T2_TurnTaking.csv")
dp3 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G2_T1_TurnTaking.csv")
dp4 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G2_T2_Synchronous.csv")
dp5 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G3_T2_TurnTaking.csv")
dp6 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G3_T3_Synchronous.csv")
dp7 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G4_T1_Synchronous.csv")
dp8 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study1_G4_T3_TurnTaking.csv")
dp9 = read.csv("~/Documents/Aarhus/3. Semester/2. Experimental Methods 3/Assignments/Assignment 4/pp/pp_Study2_G10_T1_Conversation.csv")

#Not good - get rid of data 
pResp1 = ggplot(dp1)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp1

#Not good - get rid of data 
pResp2 = ggplot(dp2)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp2

#Not good - get rid of data 
pResp3 = ggplot(dp3)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp3

#Not good - get rid of 
pResp4 = ggplot(dp4)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp4

#Keep this data 
pResp5 = ggplot(dp5)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp5

pHR5 = ggplot(dp5)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR5

#Keep this data 
pResp6 = ggplot(dp6)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp6

pHR6 = ggplot(dp6)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR6

#Not good - get rid of data 
pResp7 = ggplot(dp7)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp7

#Keep data 
pResp8 = ggplot(dp8)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp8

pHR8 = ggplot(dp8)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR8

#Keep data
pResp9 = ggplot(dp9)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp9

pHR9 = ggplot(dp9)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR9

dpHR10 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T2_TurnTaking.csv")
pHR10 = ggplot(dpHR10)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR10

dpResp10 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T2_TurnTaking.csv")
pResp10 = ggplot(dpResp10)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp10


dpHR11 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T3_Synchronous.csv")
pHR11 = ggplot(dpHR11)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR11

dpResp11 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G10_T3_Synchronous.csv")
pResp11 = ggplot(dpResp11)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp11

dpHR12 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T1_Synchronous.csv")
pHR12 = ggplot(dpHR12)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR12

dpResp12 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T1_Synchronous.csv")
pResp12 = ggplot(dpResp12)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp12

dpHR13 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T2_TurnTaking.csv")
pHR13 = ggplot(dpHR13)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR13

dpResp13 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T2_TurnTaking.csv")
pResp13 = ggplot(dpResp13)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp13


dpHR14 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T3_Conversation.csv")
pHR14 = ggplot(dpHR14)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR14

dpResp14 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G5_T3_Conversation.csv")
pResp14 = ggplot(dpResp14)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp14

dpHR15 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T1_TurnTaking.csv")
pHR15 = ggplot(dpHR15)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR15

dpResp15 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T1_TurnTaking.csv")
pResp15 = ggplot(dpResp15)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp15

dpHR16 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T2_Conversation.csv")
pHR16 = ggplot(dpHR16)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR16

dpResp16 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T2_Conversation.csv")
pResp16 = ggplot(dpResp16)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp16

dpHR17 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T3_Synchronous.csv")
pHR17 = ggplot(dpHR17)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR17

dpResp17 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G6_T3_Synchronous.csv")
pResp17 = ggplot(dpResp17)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp17

dpHR18 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G7_T1_Conversation.csv")
pHR18 = ggplot(dpHR18)+
  geom_line(aes(x=time, y=HR1, colour = "blue"))+
  geom_line(aes(x=time, y=HR2))
pHR18

dpResp18 <- read.csv("~/Desktop/Experimental methods 3/Assignment 4/pp/pp_Study2_G7_T1_Conversation.csv")
pResp18 = ggplot(dpResp18)+
  geom_line(aes(x=time, y=Resp1, colour = "blue"))+
  geom_line(aes(x=time, y=Resp2))
pResp18

dp19 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G7_T2_Synchronous.csv")

pResp19 =ggplot(dp19)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp19

pHR19 =ggplot(dp19)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR19

dp20 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G7_T3_TurnTaking.csv")

pResp20 =ggplot(dp20)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp20

pHR20 =ggplot(dp20)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR20

dp21 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G8_T1_TurnTaking.csv")

pResp21 =ggplot(dp21)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp21

pHR21 =ggplot(dp21)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR21

dp22 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G8_T2_Synchronous.csv")

pResp22 =ggplot(dp22)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp22

pHR22 =ggplot(dp22)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR22

dp23 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G8_T3_Conversation.csv")

pResp23 =ggplot(dp23)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp23

pHR23 =ggplot(dp23)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR23

dp24 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G9_T1_Synchronous.csv")

pResp24 =ggplot(dp24)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp24

pHR24 =ggplot(dp24)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR24

dp25 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G9_T2_Conversation.csv")

pResp25 =ggplot(dp25)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp25

pHR25 =ggplot(dp25)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR25

dp26 = read_csv("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/pp_Study2_G9_T3_TurnTaking.csv")

pResp26 =ggplot(dp26)+
  geom_line(aes(x=time, y=Resp1, colour = "red"))+
  geom_line(aes(x=time, y=Resp2))
pResp26

pHR26 =ggplot(dp26)+
  geom_line(aes(x=time, y=HR1, colour = "red"))+
  geom_line(aes(x=time, y=HR2))
pHR26


```
### Systematically pre-process the data
- Loop through all the files (either with a loop or with a function), check which files should be excluded, if any, and save the pre-processed time-series. Tip: plot and visually inspect the data to figure out which should be excluded.
- Run crqa on all the pre-processed time-series and save the output (don't forget to add columns with study, group, condition and trial). Tip: remember to first assess optimal parameters (dimensions, delay, radius) across all timeseries. Tip: it will often fail, just take whatever parameters you get, select optimal across timeseries parameters and run crqa on all timeseries with those. Tip: double check the rr. When I ran the loop, I got very low rr, so I adjusted the radius until the average of rr across all pairs was approx. 4%.
```{r}
setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp")


#CRQA loop
files1 <- list.files(path = ("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp"), pattern = "pp", recursive = TRUE)
files1

study = NULL
pair = NULL
trial = NULL
condition = NULL
HRdelay = NULL
HRembed = NULL
HRradius = NULL
Respdelay = NULL
Respembed = NULL
Respradius = NULL
n = 1

for (i in files1){
  print(i)
  file = read.csv(i)
  # study[n] = substr(i,start = 9, stop = 10)
  # pair[n] = substr(i,start = 11, stop = 13)
  # trial[n] = substr(i,start = 15, stop = 16)
  # condition[n] = substr(i, start = 17, stop = 20)
  study[n] = str_extract(i,"Study+\\d+")
  pair[n] = str_extract(i,"G+\\d+")
  trial[n] = str_extract(i,"T+\\d+")
  condition[n] = substr(i, start = 17, stop = 20)
  
  par = list(lgM =  30, steps = seq(1, 6, 1),  radiusspan = 100, radiussample = 100, normalize = 0,  rescale = 0,  mindiagline = 2,  minvertline = 2,  tw = 0,  whiteline = FALSE,  recpt = FALSE,  fnnpercent = 10,  typeami = "mindip")
  #widen the range of the recurrence (3-5 instead)
  ans = try(optimizeParam(file$HR1S, file$HR2S, par, min.rec = 2, max.rec = 8))
  #if the length of values calculated from ans is less than 2, then write NA
  if(length(ans) < 2) {
    HRdelay[n] = NA
    HRembed[n] = NA
    HRradius[n] = NA
  }
  # If greater than 2 use the calculated values
  else {
    HRdelay[n]=ans$delay
    HRembed[n]=ans$emddim
    HRradius[n]=ans$radius
  }
  ans = try(optimizeParam(file$Resp1S, file$Resp2S, par, min.rec = 2, max.rec = 8))
  #if the length of values calculated from ans is less than 2, then write NA
  if(length(ans) < 2) {
    Respdelay[n] = NA
    Respembed[n] = NA
    Respradius[n] = NA
  }
  # If greater than 2 use the calculated values
  else {
    Respdelay[n]=ans$delay
    Respembed[n]=ans$emddim
    Respradius[n]=ans$radius
  }
  #making sure that the next is saved in the new row
  n=n+1
}

test =data.frame(study, pair, trial,condition,HRdelay,HRembed, HRradius, Respdelay, Respembed, Respradius)

#median for Heart rate
delayHR= round(median(HRdelay, na.rm=T))#- sd(HRdelay, na.rm = T))
embedHR= round(median(HRembed, na.rm=T)) #- sd(HRembed, na.rm = T))
radiusHR=round(median(HRradius, na.rm=T),2) + 1.2


#median for Resp
delayResp= round(median(Respdelay, na.rm=T)) #- sd(Respdelay, na.rm = T))
embedResp= round(median(Respembed, na.rm=T))# - sd(Respembed, na.rm = T))
radiusResp=round(median(Respradius, na.rm=T),2)


```


```{r}
setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp")

#HR data
RR_HR = NULL
DET_HR <- NULL
NRLINE_HR <- NULL
maxL_HR <- NULL
L_HR <- NULL
ENTR_HR <- NULL
LAM_HR <- NULL
TT_HR <- NULL
#Resp data
RR_Resp = NULL
DET_Resp <- NULL
NRLINE_Resp <- NULL
maxL_Resp <- NULL
L_Resp <- NULL
ENTR_Resp <- NULL
LAM_Resp <- NULL
TT_Resp <- NULL

n = 1
for (i in files1){
  file = read.csv(i)
  print(i)
  # the crqa with the calculated delay and embed, and a radius of 3 to get the optimal recurrence percentage at approx 5-10%
  resultsHR = try(crqa(file$HR1S,file$HR2S, delay=delayHR, embed=embedHR, radius=radiusHR,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(resultsHR) > 1) {
    RR_HR[n] <- resultsHR$RR
    DET_HR[n] <- resultsHR$DET
    maxL_HR[n] <- resultsHR$maxL
    L_HR[n] <- resultsHR$L
    ENTR_HR[n] <- resultsHR$ENTR
    LAM_HR[n] <- resultsHR$LAM
    TT_HR[n] <- resultsHR$TT
  }
  resultsResp = try(crqa(file$Resp1S,file$Resp2S, delay=delayResp, embed=embedResp, radius=radiusResp,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(resultsResp) > 1) {
    RR_Resp[n] <- resultsResp$RR
    DET_Resp[n] <- resultsResp$DET
    maxL_Resp[n] <- resultsResp$maxL
    L_Resp[n] <- resultsResp$L
    ENTR_Resp[n] <- resultsResp$ENTR
    LAM_Resp[n] <- resultsResp$LAM
    TT_Resp[n] <- resultsResp$TT
  }
  n= n+1
}

#final dataframe
final_data =data.frame(study, pair, trial,condition,RR_HR,DET_HR, maxL_HR, L_HR,ENTR_HR, LAM_HR, TT_HR ,RR_Resp, DET_Resp, maxL_Resp, L_Resp, ENTR_Resp, LAM_Resp, TT_Resp)
```
### Creating controls: shuffled controls
 - loop through all pairs and conditions
 - shuffle the timeseries (take a timeseries and rearrange its values in a random order). Tip check the sample() function
 - run crqa and save the output. NB. which delay, embed, radius parameters should you use?
 - statistically compare the crqa indexes in real and shuffled pairs
 - xshu = sample(x, lenght)
 
```{r}
setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp")

for (i in files1){
  file = read.csv(i)
  print(i)
  shuffleddata = file[sample(1:nrow(file),2401,replace = T ), ] 
    file2 = paste0("shu_", i)
  write_csv(shuffleddata, file.path("shu", file2))
}

setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/shu")
files2 <- list.files(path = ("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/shu"), pattern = "shu", recursive = TRUE)
files2


#HR data for shuffled
RR_HR1 = NULL
DET_HR1 <- NULL
maxL_HR1 <- NULL
L_HR1 <- NULL
ENTR_HR1 <- NULL
LAM_HR1 <- NULL
TT_HR1 <- NULL
#Resp data
RR_Resp1 = NULL
DET_Resp1 <- NULL
maxL_Resp1 <- NULL
L_Resp1 <- NULL
ENTR_Resp1 <- NULL
LAM_Resp1 <- NULL
TT_Resp1 <- NULL

n = 1
for (i in files2){
  file = read.csv(i)
  print(i)
  # the crqa with the calculated delay and embed, and a radius of 3 to get the optimal recurrence percentage at approx 5-10%
  resultsHR1 = try(crqa(file$HR1S,file$HR2S, delay=delayHR, embed=embedHR, radius=radiusHR,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(resultsHR1) > 1) {
    RR_HR1[n] <- resultsHR1$RR
    DET_HR1[n] <- resultsHR1$DET
    maxL_HR1[n] <- resultsHR1$maxL
    L_HR1[n] <- resultsHR1$L
    ENTR_HR1[n] <- resultsHR1$ENTR
    LAM_HR1[n] <- resultsHR1$LAM
    TT_HR1[n] <- resultsHR1$TT
  }
  resultsResp1 = try(crqa(file$Resp1S,file$Resp2S, delay=delayResp, embed=embedResp, radius=radiusResp,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(resultsResp1) > 1) {
    RR_Resp1[n] <- resultsResp1$RR
    DET_Resp1[n] <- resultsResp1$DET
    maxL_Resp1[n] <- resultsResp1$maxL
    L_Resp1[n] <- resultsResp1$L
    ENTR_Resp1[n] <- resultsResp1$ENTR
    LAM_Resp1[n] <- resultsResp1$LAM
    TT_Resp1[n] <- resultsResp1$TT
  }
  n= n+1
}

final_data1 =data.frame(study, pair, trial,condition,RR_HR=RR_HR1,DET_HR = DET_HR1,maxL_HR= maxL_HR1, L_HR =L_HR1,ENTR_HR = ENTR_HR1, LAM_HR =LAM_HR1, TT_HR = TT_HR1 ,RR_Resp =RR_Resp1,DET_Resp= DET_Resp1, maxL_Resp =maxL_Resp1,L_Resp= L_Resp1, ENTR_Resp =ENTR_Resp1, LAM_Resp =LAM_Resp1, TT_Resp =TT_Resp1)

final_data1$type = "shuffled"
final_data$type = "real"

newdata = rbind(final_data, final_data1)


real_vs_shuffledHR = lmer(RR_HR ~ type + (1|pair), newdata)
summary(real_vs_shuffled)
#significant = different 

real_vs_shuffledResp = lmer(RR_Resp ~ type + (1|pair), newdata)
summary(real_vs_shuffledResp)

#add a column to the fana
```
 
 
### TRICKY! Creating controls: surrogate pair controls
 - Per each real pair, identify at least one surrogate pair (matching one of the participants, with somebody doing the same task, but in a different pair). Tip: Celine will share a commented script
 - Run crqa on all the surrogate pairs and save the output. NB. which delay, embed, radius parameters should you use?
 - Test whether crqa shows a difference between real and surrogate pairs
```{r}
#surrogate script from celine
setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp")

final_data$condition[which(final_data$condition=="_Tur")] <- "Turn"
final_data$condition[which(final_data$condition=="_Con")] <- "Conv"
final_data$condition[which(final_data$condition=="_Syn")] <- "Sync"


prepro_filelist = list.files(pattern = "pp_Study2") # load the preprocessed files

num=str_extract_all(prepro_filelist, "\\d+") #extract all numbers in the file name (i.e Study?_G?_T?)

df= as.data.frame(t(as.data.frame(num))) #create a transposed (i.e. "tipped") dataframe of these number
rownames(df)<-NULL #Remove rownames.

pairs=as.numeric(unique(as.numeric(as.character(df$V2)))) #create pairs, i.e. unique group numbers

SurrogateList <- expand.grid(a = pairs, b = pairs) # expand.grid combines all possible combinations of a given vector
SurrogateList = subset(SurrogateList, a != b) # Ensure that pair 1 (i.e group1) doesn't get paired with it self.


# Per each condition
for (co in c("Sync","Turn", "Conv")){ #loop through a list of condintions 
  for (i in 1:nrow(SurrogateList)){ #and loop through all the pairings of the surrogate list
    
  #start recreate filenames from the working directory based on the matching from the surrogatelist
  f1=paste0('pp_Study2_G',SurrogateList$a[i]) #Insert group numeber from Surrogatelist 
  f2=paste0('pp_Study2_G',SurrogateList$b[i])
  
  file1=prepro_filelist[grepl(f1,prepro_filelist) & grepl(co,prepro_filelist)] #insert conditions as well.
  file2=prepro_filelist[grepl(f2,prepro_filelist) & grepl(co,prepro_filelist)]
  
  d1=read.csv(file1) #Load the recreated filenames
  
  d2=read.csv(file2) 
  
  #Ensure both dataframes have the same length
  if (nrow(d1)>nrow(d2)){
    d1=d1[1:nrow(d2),]
  } else {
    d2=d2[1:nrow(d1),]
  }
  
  #creation of surrogate files. (Riccardo's idea of having fun and completely overdoing it ;-) ) 
  #per each 2 pairs there are 4 possible surrogate pairs: 
  #pair1person1 w pair2person2; 
  d3=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp2,HR1=d1$HR1,HR2=d2$HR2,Resp1S=d1$Resp1,Resp2S=d2$Resp2,HR1S=d1$HR1,HR2S=d2$HR2)
  #pair1person2 w pair2person2;
  d4=data.frame(Resp1=d2$Resp1,Resp2=d1$Resp2,HR1=d2$HR1,HR2=d1$HR2,Resp1S=d2$Resp1,Resp2S=d1$Resp2,HR1S=d2$HR1,HR2S=d1$HR2)
  
  #pair1person1 w pair2person1;
  d5=data.frame(Resp1=d1$Resp1,Resp2=d2$Resp1,HR1=d1$HR1,HR2=d2$HR1,Resp1S=d1$Resp1,Resp2S=d2$Resp1,HR1S=d1$HR1,HR2S=d2$HR1)
  #pair1person2 w pair2person2
  d6=data.frame(Resp1=d2$Resp2,Resp2=d1$Resp2,HR1=d2$HR2,HR2=d1$HR2,Resp1S=d2$Resp2,Resp2S=d1$Resp2,HR1S=d2$HR2,HR2S=d1$HR2)
  
  #Make filenames that tells which pairs have been mixed
  file3=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"a_",co,".csv")
  file4=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"b_",co,".csv")
  file5=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"c_",co,".csv")
  file6=paste0("Study2_SurrogateGroup",SurrogateList$a[i],"&",SurrogateList$b[i],"d_",co,".csv")
  
  #write csv files
  write_csv(d3,file.path("surrogate",file3))
  write_csv(d4,file.path("surrogate",file4))
  write_csv(d5,file.path("surrogate",file5))
  write_csv(d6,file.path("surrogate",file6))
  }
}




```

```{r}

setwd("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/surrogate")
files3 <- list.files(path = ("~/Documents/Cognitive Science 3. semester/Experimental methods 3/Rstudio/Assignment 4/pp/surrogate"), pattern = "Study2", recursive = TRUE)
files3


#HR data for shuffled
RR_HR2 = NULL
DET_HR2 <- NULL
maxL_HR2 <- NULL
L_HR2 <- NULL
ENTR_HR2 <- NULL
LAM_HR2 <- NULL
TT_HR2 <- NULL
#Resp data
RR_Resp2 = NULL
DET_Resp2 <- NULL
maxL_Resp2 <- NULL
L_Resp2 <- NULL
ENTR_Resp2 <- NULL
LAM_Resp2 <- NULL
TT_Resp2 <- NULL

n = 1
for (i in files3){
  file = read.csv(i)
  print(i)
  # the crqa with the calculated delay and embed, and a radius of 3 to get the optimal recurrence percentage at approx 5-10%
  resultsHR2 = try(crqa(file$HR1S,file$HR2S, delay=delayHR, embed=embedHR, radius=radiusHR,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(resultsHR2) > 1) {
    RR_HR2[n] <- resultsHR2$RR
    DET_HR2[n] <- resultsHR2$DET
    maxL_HR2[n] <- resultsHR2$maxL
    L_HR2[n] <- resultsHR2$L
    ENTR_HR2[n] <- resultsHR2$ENTR
    LAM_HR2[n] <- resultsHR2$LAM
    TT_HR2[n] <- resultsHR2$TT
  }
  resultsResp2 = try(crqa(file$Resp1S,file$Resp2S, delay=delayResp, embed=embedResp, radius=radiusResp,normalize=0,rescale=0,mindiagline = 2,minvertline = 2))
  #if the length of values calculated from results is greater than 1, then use the calculated values
  if (length(resultsResp2) > 1) {
    RR_Resp2[n] <- resultsResp2$RR
    DET_Resp2[n] <- resultsResp2$DET
    maxL_Resp2[n] <- resultsResp2$maxL
    L_Resp2[n] <- resultsResp2$L
    ENTR_Resp2[n] <- resultsResp2$ENTR
    LAM_Resp2[n] <- resultsResp2$LAM
    TT_Resp2[n] <- resultsResp2$TT
  }
  n= n+1
}

final_data2 =data.frame(RR_HR=RR_HR2,DET_HR = DET_HR2,maxL_HR= maxL_HR2, L_HR =L_HR2,ENTR_HR = ENTR_HR2, LAM_HR =LAM_HR2, TT_HR = TT_HR2 ,RR_Resp =RR_Resp2,DET_Resp= DET_Resp2, maxL_Resp =maxL_Resp2,L_Resp= L_Resp2, ENTR_Resp =ENTR_Resp2, LAM_Resp =LAM_Resp2, TT_Resp =TT_Resp2)


#deleting columns from the real data
real_data = final_data
real_data$study = NULL
real_data$pair = NULL
real_data$condition= NULL
real_data$trial = NULL

final_data2$type = "surrogate"

newdata1 = rbind(real_data, final_data2)

real_vs_surrogateHR = lm(RR_HR ~ type, newdata1)
summary(real_vs_surrogateHR)
#significant = different 

real_vs_surrogateResp = lm(RR_Resp ~ type, newdata1)
summary(real_vs_surrogateResp)


```
### Testing effects of conditions
 - make a (probably underpowered) mixed model testing effects of the different conditions on heart rate and respiration coordination
 - N.B: would it make sense to include surrogate pairs? and if so how? what would that tell you?
```{r}

final_data$trial = as.numeric(final_data$trial)


diff_conHR = lmer(RR_HR~ condition+ (1|study) + (1+ trial|pair), final_data)
summary(diff_conHR)

diff_conResp = lmer(RR_Resp~ condition+ (1|study) + (1+trial|pair), final_data)
summary(diff_conResp)


```

### hypothesis:Effects of respiration coordination on heart rate coordination
 - describe how you would test those.
 - Optional: run the models and report them
